The Dynamics of Anthropomorphic Effects in Human-Robot Interaction

### 2014-03-21 paper.tex
Line wrapping at 80 char + bring back roomba perception fig

\usepackage{graphicx}
\graphicspath{{figs/}}
\usepackage{tikz}
%\usepackage{caption}
%\usepackage{subcaption}
%\usepackage{float}
%\floatstyle{plaintop}
\usetikzlibrary{shapes}
%
\usepackage{amsmath} 
\usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
\usepackage{color, soul}
\usepackage{url}
\usepackage{multirow}
\usepackage{array}
\usepackage{fixltx2e}
\usepackage{textcomp}

@@

%The main goal of this study was to explore how households live with a vacuum cleaning robot over an extended period of time. We wanted to analyze how the usage of and interaction with the robot evolve over time, and to identify adoption phases as well as factors that impact on the long-term acceptance of service robots in homes.

From the Roomba study, we can report on several findings related to anthropomorphism and the anthropomorphic design of robots.
%The gathered data for these aspects include qualitative observations, semi-structured interviews, and also quantitative assessments of various aspects of human-robot interaction.
The Roomba study is one of very few long-term human-robot interaction studies in the domestic environment.

@@

In our long-term study of the Roomba in homes, we found that it was mostly during the initial phase with the robot that some (rather few) people tended to interact in an anthropomorphic or human-social way with their vacuum cleaning robot.
Overall, our data suggest that the anthropomorphic effect and the social impact of a functional robot on a household is rather marginal.
%There were just few instances of people anthropomorphizing the robot and interacting with it in a human-social way. More importantly, overall, these anthropomorphic interactions with the robot did not last but seemed to be somehow related to the novelty of the robot.

We categorized several interactions and actions with the robot as anthropomorphic or human-social (in a broader sense). 
%Examples of what we categorized as somehow social and anthropomorphic (human-like) interactions with the robot include \textit{watching} the robot for fun, giving a \textit{demo} of the robot to guests / visitors, taking a \textit{photo/video} of the robot, \textit{personalizing} the robot, giving a \textit{name} to the robot, \textit{talking} to the robot using direct speech, \textit{playing} with the robot, and making \textit{adjustments} to the home due to the robot. 
Figure \ref{fig:roomba-activities}) shows how many of the 9 households engaged in these activities during the 6 months when the study was conducted.

\begin{figure}
    \centering
    \includegraphics[width=0.5\columnwidth]{roomba-activities.png}
    \caption{Activities with the robot over time, n = 9 households. This bar chart shows how many of the 9 households carried out different kinds of anthropomorphic / social activities with and related to the robot at 3 time points during 6 months. Activities include \texttt{watching} the robot for fun, giving a \texttt{demo} of the robot to guests / visitors, \texttt{personalizing} the robot, giving a \texttt{name} to the robot, \texttt{talking} to the robot, \texttt{playing} with the robot. Participants were asked to report on these activities at 3 time points after the robot was deployed in their home: after 2 weeks, after 2 months, and after 6 months.}
    \label{fig:roomba-activities}
\end{figure}

@@

%\paragraph*{Taking photos / videos of the robot.} How far taking a photo or video of a robot can be seen as an anthropomorphic interaction is questionable. However, it reflects the fact that the robot is observed by the user who for some reason is motivated to record the event. The number of households that took a photo or video of the robot decreased over time from 5 households in the beginning to 1 household after 6 months.

@@

%\paragraph*{Making adjustments to the home due to the robot.} To help Roomba clean, \cite{sung_housewives_2008} found in an online survey that about half of the 379 participants made physical modifications to their homes. This process is referred to as \textit{roombarization}, and it seems to be a unique aspect of adopting a domestic service robot. In our study, almost all the households made adjustments when they used the robot for the first time in front of us. However, during the first two weeks with the robot, only 3 households reported to have adjusted something due to the robot, and after 6 months it was only 1 household that still engaged in this activity. This decrease is, however, logical since after some time most people had done all the necessary adjustments to enable the robot show a good performance.

%\paragraph*{Emerging anthropomorphic interactions.} Another, from our side more unexpected interaction that we would describe as somehow anthropomorphic, is what we call \textit{``collaborative cleaning''} with the robot. In the initial phase of the study, some people spontaneously helped the robot to clean by either picking up crumbs from the floor themselves or pushing it right in front of the robot. Another way of collaborating with the robot we observed was to help it go everywhere by spontaneously putting up the trash bin or the carpet in the bathroom (things that are not just lying around and need to be cleaned up). However, the activity of collaborative cleaning faded out soon, and we suspect it was encouraged by the novelty of the robot.\\

@@

%\begin{figure*}[htb]
%\centering
%
%\begin{tikzpicture}[scale=0.5, y = 2cm]
%%axis
%\draw (0,0) -- coordinate (x axis mid) (20,0);
%\draw (0,0) -- coordinate (y axis mid) (0,7);
%%ticks
%\foreach \x in {0,4,...,20}
%    \draw (\x,1pt) -- (\x,-3pt);
%\foreach \y in {1,...,7} 
%    \draw[black!20] (20,\y) -- (-3pt,\y);
%\foreach \y in {1,...,7} 
%    \draw (-3pt, \y) node[anchor=east] {\y}; 
%%labels
%\node[below=0.8cm] at (x axis mid) {People's perception ($\sigma=15$)};
%\node at (2, -0.4) {\tiny expectations};
%\node[align=center] at (6, -0.4) {\tiny after\\ \tiny first usage};
%\node[align=center] at (10, -0.4) {\tiny after\\ \tiny 2 weeks};
%\node[align=center] at (14, -0.4) {\tiny after\\ \tiny 2 months};
%\node[align=center] at (18, -0.4) {\tiny after\\ \tiny 6 months};
%
%\draw[dashed, black!50] (6,0) -- (6,7) node[sloped, below, pos=0.25] {\tiny novelty effect};
%
%\draw[ultra thick, red!50] (2,5.6) -- (6,6.1) -- (10,6.3) -- (14,6.3) -- (18,6.1) node at (15,6.5) {\tiny \it ease of use};
%\draw[ultra thick, purple] (2,4.9) -- (6,5.2) -- (10,5.1) -- (14,4.5) -- (18,4.2) node at (3,4.7) {\tiny \it usefulness};
%\draw[ultra thick, blue!50] (2,4) -- (6,4.1) -- (10,4.6) -- (14,4.1) -- (18,4.3) node at (16,4) {\tiny \it fun};
%\draw[ultra thick, green!50] (2,3.9) -- (6,4.1) -- (10,4.2) -- (14,3.8) -- (18,3.3) node at (19,3.5) {\tiny \it intelligence};
%\draw[ultra thick, orange] (2,3.5) -- (6,4.2) -- (10,4) -- (14,3.2) -- (18,2.8) node at (14,2.7) {\tiny \it impact};
%
%\draw[ultra thick, dashed] (2,5.2) -- (6,5.2) -- (10,5) -- (14,4.9) -- (18,4.5) node at (14,5.3) {\tiny \it overall};
%
%\end{tikzpicture}
%
%\caption{User perception of the Roomba robot over a 6 month period
%($\sigma=15$).}
%
%\label{roomba-perception}
%\end{figure*}

@@

\begin{figure}
    \centering
    \includegraphics[width=0.5\columnwidth]{roomba-perception.pdf}
    \caption{People's perception of the robot over time, n = 15 participants. This line chart shows how people's perception of the robot evolves over time. At 5 time points people were asked to rate the robot according to the following aspects: the perceived \texttt{intelligence} of the robot, the perceived \texttt{usefulness}, the perceived \texttt{ease of use}, how much \texttt{fun} it was to use the robot, how far they found the robot had an \texttt{impact} on their household, and their {overall} impression of it.}
    \label{fig:roomba-perception}
\end{figure}

@@

\begin{enumerate}
	\item People engage in a more social and human-like way with a robotic pet compared to a functional robot and a tablet computer, therefore there will be a higher amount of anthropomorphic language in the AIBO and Roomba forum, compared to the iPad forum.
	\item As anthropomorphism can be seen as a certain kind of emotional / social relation to an artifact, there will be more anthropomorphic language in forum posts in which the author describes a relation to the device, compared to discussing technological aspects or how the device is used.
\end{enumerate}

@@

\begin{itemize}
	\item \textit{Life-likeness}, uch as being alive or having (parts of) a body, \eg \textit{``she can be reborn since you copied her memory stick''};
	\item \textit{Emotional states} or the technology having a feeling, \eg \textit{``oh no, poor AIBO''};
	\item \textit{Gender, personality} or the technology having an \textit{intention}, \eg Roomba \textit{``seems to like to hang around under the sofa''};
	\item When the author gave the technology a \textit{name}, \eg \textit{``whenever I show Java his pink ball''};
	\item Being \textit{socially integrated}, such as being a family member, \eg \textit{``I'm considering adding a Roomba to the family''};
	\item \textit{Metaphorical ways}, \eg Roomba \textit{``sings its victory song when it finishes and docks''}.
\end{itemize}

>>>>

\usepackage{graphicx}
\graphicspath{{figs/}}
\usepackage{tikz}
\usetikzlibrary{shapes}
%
\usepackage{amsmath} 
\usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
\usepackage{color, soul}
\usepackage{url}
\usepackage{multirow}
\usepackage{array}
\usepackage{fixltx2e}
\usepackage{textcomp}

@@

%The main goal of this study was to explore how households live with a vacuum
%cleaning robot over an extended period of time. We wanted to analyze how the
%usage of and interaction with the robot evolve over time, and to identify
%adoption phases as well as factors that impact on the long-term acceptance of
%service robots in homes.

From the Roomba study, we can report on several findings related to
anthropomorphism and the anthropomorphic design of robots.
%The gathered data for these aspects include qualitative observations,
%semi-structured interviews, and also quantitative assessments of various
%aspects of human-robot interaction.
The Roomba study is one of very few long-term human-robot interaction studies in
the domestic environment.

@@

In our long-term study of the Roomba in homes, we found that it was mostly
during the initial phase with the robot that some (rather few) people tended to
interact in an anthropomorphic or human-social way with their vacuum cleaning
robot.  Overall, our data suggest that the anthropomorphic effect and the social
impact of a functional robot on a household is rather marginal.
%There were just few instances of people anthropomorphizing the robot and
%interacting with it in a human-social way. More importantly, overall, these
%anthropomorphic interactions with the robot did not last but seemed to be
%somehow related to the novelty of the robot.

We categorized several interactions and actions with the robot as
anthropomorphic or human-social (in a broader sense). 
%Examples of what we categorized as somehow social and anthropomorphic
%(human-like) interactions with the robot include \textit{watching} the robot
%for fun, giving a \textit{demo} of the robot to guests / visitors, taking a
%\textit{photo/video} of the robot, \textit{personalizing} the robot, giving a
%\textit{name} to the robot, \textit{talking} to the robot using direct speech,
%\textit{playing} with the robot, and making \textit{adjustments} to the home
%due to the robot. 
Figure \ref{fig:roomba-activities}) shows how many of the 9 households engaged
in these activities during the 6 months when the study was conducted.

\begin{figure}
    \centering
    \includegraphics[width=0.5\columnwidth]{roomba-activities.png}
    \caption{Activities with the robot over time, n = 9 households. This bar
        chart shows how many of the 9 households carried out different kinds of
        anthropomorphic / social activities with and related to the robot at 3
        time points during 6 months. Activities include \texttt{watching} the
        robot for fun, giving a \texttt{demo} of the robot to guests / visitors,
        \texttt{personalizing} the robot, giving a \texttt{name} to the robot,
        \texttt{talking} to the robot, \texttt{playing} with the robot.
        Participants were asked to report on these activities at 3 time points
        after the robot was deployed in their home: after 2 weeks, after 2 months,
        and after 6 months.}

\label{fig:roomba-activities}
\end{figure}

@@

%\paragraph*{Taking photos / videos of the robot.} How far taking a photo or
%video of a robot can be seen as an anthropomorphic interaction is questionable.
%However, it reflects the fact that the robot is observed by the user who for
%some reason is motivated to record the event. The number of households that
%took a photo or video of the robot decreased over time from 5 households in the
%beginning to 1 household after 6 months.

@@

%\paragraph*{Making adjustments to the home due to the robot.} To help Roomba
%clean, \cite{sung_housewives_2008} found in an online survey that about half of
%the 379 participants made physical modifications to their homes. This process
%is referred to as \textit{roombarization}, and it seems to be a unique aspect
%of adopting a domestic service robot. In our study, almost all the households
%made adjustments when they used the robot for the first time in front of us.
%However, during the first two weeks with the robot, only 3 households reported
%to have adjusted something due to the robot, and after 6 months it was only 1
%household that still engaged in this activity. This decrease is, however,
%logical since after some time most people had done all the necessary
%adjustments to enable the robot show a good performance.

%\paragraph*{Emerging anthropomorphic interactions.} Another, from our side more
%unexpected interaction that we would describe as somehow anthropomorphic, is
%what we call \textit{``collaborative cleaning''} with the robot. In the initial
%phase of the study, some people spontaneously helped the robot to clean by
%either picking up crumbs from the floor themselves or pushing it right in front
%of the robot. Another way of collaborating with the robot we observed was to
%help it go everywhere by spontaneously putting up the trash bin or the carpet
%in the bathroom (things that are not just lying around and need to be cleaned
%up). However, the activity of collaborative cleaning faded out soon, and we
%suspect it was encouraged by the novelty of the robot.\\

@@

@@

\begin{figure}
    \centering
    \includegraphics[width=0.5\columnwidth]{roomba-perception.pdf}

    \begin{tikzpicture}[scale=0.5, y = 2cm]
    %axis
    \draw (0,0) -- coordinate (x axis mid) (20,0);
    \draw (0,0) -- coordinate (y axis mid) (0,7);
    %ticks
    \foreach \x in {0,4,...,20}
        \draw (\x,1pt) -- (\x,-3pt);
    \foreach \y in {1,...,7} 
        \draw[black!20] (20,\y) -- (-3pt,\y);
    \foreach \y in {1,...,7} 
        \draw (-3pt, \y) node[anchor=east] {\y}; 
    %labels
    \node[below=0.8cm] at (x axis mid) {People's perception ($\sigma=15$)};
    \node at (2, -0.4) {\tiny expectations};
    \node[align=center] at (6, -0.4) {\tiny after\\ \tiny first usage};
    \node[align=center] at (10, -0.4) {\tiny after\\ \tiny 2 weeks};
    \node[align=center] at (14, -0.4) {\tiny after\\ \tiny 2 months};
    \node[align=center] at (18, -0.4) {\tiny after\\ \tiny 6 months};

    \draw[dashed, black!50] (6,0) -- (6,7) node[sloped, below, pos=0.25] {\tiny novelty effect};

    \draw[ultra thick, red!50] (2,5.6) -- (6,6.1) -- (10,6.3) -- (14,6.3) -- (18,6.1) node at (15,6.5) {\tiny \it ease of use};
    \draw[ultra thick, purple] (2,4.9) -- (6,5.2) -- (10,5.1) -- (14,4.5) -- (18,4.2) node at (3,4.7) {\tiny \it usefulness};
    \draw[ultra thick, blue!50] (2,4) -- (6,4.1) -- (10,4.6) -- (14,4.1) -- (18,4.3) node at (16,4) {\tiny \it fun};
    \draw[ultra thick, green!50] (2,3.9) -- (6,4.1) -- (10,4.2) -- (14,3.8) -- (18,3.3) node at (19,3.5) {\tiny \it intelligence};
    \draw[ultra thick, orange] (2,3.5) -- (6,4.2) -- (10,4) -- (14,3.2) -- (18,2.8) node at (14,2.7) {\tiny \it impact};

    \draw[ultra thick, dashed] (2,5.2) -- (6,5.2) -- (10,5) -- (14,4.9) -- (18,4.5) node at (14,5.3) {\tiny \it overall};

    \caption{People's perception of the robot over time, n = 15 participants.
    This line chart shows how people's perception of the robot evolves over
    time. At 5 time points people were asked to rate the robot according to the
    following aspects: the perceived \texttt{intelligence} of the robot, the
    perceived \texttt{usefulness}, the perceived \texttt{ease of use}, how much
    \texttt{fun} it was to use the robot, how far they found the robot had an
    \texttt{impact} on their household, and their {overall} impression of it.}

    \label{fig:roomba-perception}
\end{figure}

@@

    	\item People engage in a more social and human-like way with a robotic pet compared to a functional robot and a tablet computer, therefore there will be a higher amount of anthropomorphic language in the AIBO and Roomba forum, compared to the iPad forum.

    	\item As anthropomorphism can be seen as a certain kind of emotional / social relation to an artifact, there will be more anthropomorphic language in forum posts in which the author describes a relation to the device, compared to discussing technological aspects or how the device is used.
\end{enumerate}

@@

\begin{itemize}
    \item \textit{Life-likeness}, uch as being alive or having (parts of) a
        body, \eg \textit{``she can be reborn since you copied her memory
        stick''};

    \item \textit{Emotional states, empathy} or the technology having a
        \textit{feeling}, \eg \textit{``oh no, poor AIBO''};

    \item \textit{Gender, personality} or the technology having an
        \textit{intention}, \eg Roomba \textit{``seems to like to hang around
        under the sofa''};

    \item When the author gave the technology a \textit{name}, \eg
        \textit{``whenever I show Java his pink ball''};

    \item Being \textit{socially integrated}, such as being a family member, \eg
        \textit{``I'm considering adding a Roomba to the family''};

    \item \textit{Metaphorical ways}, \eg Roomba \textit{``sings its victory
        song when it finishes and docks''}.

\end{itemize}
